---
title: "Clustering and classification"
author: "jouni hirvonen"
date: "20 11 2019"
output: html_document
### Install required packages

First, we install the required packages for the analysis: tidyr, ggplot2, GGally, gridExtra (for multiple plots)

 ```{r, results="hide", message=FALSE, warning=FALSE}
install.packages("tidyr", repos = "http://cloud.r-project.org")
install.packages("ggplot2", repos = "http://cloud.r-project.org")
install.packages("GGally", repos = "http://cloud.r-project.org")
install.packages("gridExtra", repos = "http://cloud.r-project.org")
 ```
and make the packages available

 ```{r, results="hide", message=FALSE}
library(tidyr); library(dplyr); library(ggplot2); library(gridExtra)
```


Read data from data :

```{r}
library(MASS)

# load the data
data("Boston")
str(Boston)
```
Data has  14 variables and 506 observations

Glimpse at the boston data
```{r}
glimpse(Boston) 
```

Do some pairs 
```{r}
pairs(Boston)
```

And then drawing a bar plot of each variable

```{r,  message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
gather(Boston) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

Scaling data first take Boston data and summaraised it

```{r}
summary(Boston)
```

Then  scaling so we can compare data
```{r}
# center and standardize variables
boston_scaled <- scale(Boston)

# summaries of the scaled variables
summary(boston_scaled)

```
We can see all values has change in same scale

```{r}
# class of the boston_scaled object
class(boston_scaled)

# change the object to data frame
boston_scaled <- as.data.frame(boston_scaled)
```

