---
title: "Chapter6"
author: "Jouni Hirvonen"
date: "8 12 2019"
output: html_document
---
# RStudio Exercise 6

## Data Wrangling

See file `meet_and_repeat.R` in [my GitHub repository](https://github.com/jhirx/IODS-project/tree/master/data)  


## start of analysis


Read the `RATS` data in long form:

```{r, echo=FALSE, warning=FALSE}
RATSL <- read.table("data/RATSL.txt", header=TRUE)

str(RATSL)
```


```{r}

RATSL$Time <- as.numeric(RATSL$Time)
```


```{r}
RATSL$ID <- as.factor(RATSL$ID)
RATSL$Group <- as.factor(RATSL$Group)
```

### Graphic displays of `RATS` data

We begin with a graphical display of `RATS` data by plotting the points
for each subject to show the weight development of individual rats.




```{r, echo=FALSE}
library(ggplot2)
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```


```{r, echo=FALSE}
# Draw the plot
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))
```



```{r, echo=FALSE}
library(dplyr)
# Standardize the variable bprs
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()
# Glimpse the data
glimpse(RATSL)
```


```{r, echo=FALSE}
# Plot again with the standardised bprs
ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "Standardized Weight")
```




```{r, echo=FALSE}
# Number of observation periods
n <- RATSL$Time %>% unique() %>% length()
RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()
# Glimpse the data
glimpse(RATSS)
summary(RATSS)
```


```{r, echo=FALSE}
# Plot the mean profiles
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.9,0.5)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")
```


```{r, echo=FALSE}
RATSL$Time1 <- factor(RATSL$Time, levels = c(1,8,15,22,29,36,43,44,50,57,64))
ggplot(RATSL, aes(x = Time1, y = Weight, fill = Group)) +
  geom_boxplot()
```

### Summary graphs
# Create a summary data by treatment and subject with mean as the summary variable


```{r, echo=FALSE}

RATSL1 <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
# Draw a boxplot of the mean versus treatment
ggplot(RATSL1, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Days 7-64")
```



```{r, echo=FALSE}
# Get rid of the outliers
RATSL11 <- RATSL1 %>%
  filter(mean < 550)
# Draw a boxplot of the mean versus treatment
ggplot(RATSL11, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Days 7-64")
```



### Means testing
# T for test and A for Anova

```{r, echo=FALSE}
RATS <-  read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep  ="\t", header = T)
#RATS <- read.table("data/RATS.txt", header=TRUE)

#t.test(mean ~ baseline, data = RATS, var.equal = TRUE)

RATSL111 <- RATSL1 %>%
  mutate(baseline = RATS$WD1)
# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline + Group, data = RATSL111)
# Compute the analysis of variance table for the fitted model with anova()
anova(fit)
```


