---
title: "Chapter5"
author: "jhi"
date: "1 12 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RStudio Exercise 5

## Data wrangling

See file `create_human.R` in (https://github.com/jhirx/IODS-project/tree/master/data)  



## Analysis

First, load the data

```{r, warning=FALSE, results='hide'}
##human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep  =",", header = T)
human = read.table("data/human2.txt", header=TRUE, sep=",")
```
Access library `GGally`, `corrplot`  and  dplyr to be used in the analysis

```{r, warning=FALSE, results='hide'}
#install.packages("corrplot")
#install.packages("corrplot")
#install.packages("dplyr")

library(GGally)
library(corrplot)
library(dplyr)
```
1. Show a graphical overview of the data and show summaries of the variables in the data. Describe and interpret the outputs, commenting on the distributions of the variables and the relationships between them

first we summarizing the data
```{r, warning=FALSE}
summary(human)
```
Lowest life expectancy (Life.Exp) is 49 years and the highest is 83.5. 
Females obtain less education than males (Edu2.Ratio) and particpate less in the labor force (LFP.Ratio)


### Data visualization

Then, visualize the data

```{r, message=FALSE, warning=FALSE, fig.width=14, fig.height=18}
# visualize the 'human' variables
ggpairs(human)
```

Most countries in the sample are poor, given that the GNI is to the left.

Next more detail at the correlations. 

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
# compute the correlation matrix and visualize it with corrplot
cor(human) %>% corrplot
```
Negatively correlated are
Maternal mortality (Mat.Mor) and 
adolescent fertility rate (Ado.Birth)  
to male (`Edu2.Ratio`) 

Positively correlated are
gross national income (GNI) 
to female education (Edu2.Ratio)
and life expectancy (Life.Exp)
and expected education (Edu.Exp).



### Principal components analysis (non-standardized data)

2. Perform principal component analysis (PCA) on the not standardized human data. Show the variability captured by the principal components. Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables.

Principal components analysis (PCA) on the not standardized human data.

```{r, message=FALSE, warning=FALSE}
# perform principal component analysis
pca_human <- prcomp(human)
summary(pca_human)
```

First principal component (PC1) explains 99% of the variation in the data.

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

The biplot shows that the first principal component explain much of the cross-country variation, which is largely driven by gross national income (GNI).


### Principle components analysis (standardized data)
3. Standardize the variables in the human data and repeat the above analysis. Interpret the results of both analysis (with and without standardizing). Are the results different? Why or why not? Include captions (brief descriptions) in your plots where you describe the results by using not just your variable names, but the actual phenomenons they relate to.


```{r, message=FALSE, warning=FALSE}
# perform principal component analysis (with the SVD method)
human_std <- scale(human)
```

```{r, message=FALSE, warning=FALSE}
# perform principal component analysis (with the SVD method)
pca_human_std <- prcomp(human_std)
```

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
summary(pca_human_std)
```
Standardized data we can see first five principal components explain 92% of the variation in the data.
standardized data, large values of the GNI expalin most result


```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
# draw a biplot
biplot(pca_human_std, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

Positive weights gross national income (`GNI`) are
maternal mortality (Mat.Mor) and  adolescent birth rate (Ado.Birth)
and negative
secondary education (Edu2.Ratio),
expected years of education (Edu.Exp) 
and life expectancy (Life.Exp) 



### Interpretation of principal components  
Money talks and it seems to explain result



### Tea dataset from FactoMineR


5.Load the tea dataset from the package Factominer. Explore the data briefly: look at the structure and the dimensions of the data and visualize it. Then do Multiple Correspondence Analysis on the tea data (or to a certain columns of the data, it’s up to you). Interpret the results of the MCA and draw at least the variable biplot of the analysis. You can also explore other plotting options for MCA. Comment on the output of the plots.

Install package `FactoMineR`, load dataset `tea`.
explore that data:

```{r, message=FALSE, warning=FALSE, results='hide'}
install.packages('FactoMineR', repos = "http://cloud.r-project.org")
library(FactoMineR)
data(tea)
```

```{r, message=FALSE, warning=FALSE}
summary(tea)
str(tea)
```

The data includes 300 observations of 36 variables.


```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
# visualize the dataset
library(tidyr)
gather(tea) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

Next we select the variables of interest: `Tea`, `How`, `how`, `where`, `price`, `age`, and `lunch`

```{r, message=FALSE, warning=FALSE}
# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "where")
# select the 'keep_columns' to create a new dataset
tea_vars <- select(tea, one_of(keep_columns))
head(tea_vars)
 ```

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
# multiple correspondence analysis
mca <- MCA(tea_vars, graph = FALSE)
# summary of the model
summary(mca)
```

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")
```

Unpackaged tea is consumed in a tea shop
bagged tea is purchased from a chain store. 
Earl gray tea is more likely to be consumed with milk
and black tea with lemon and green tea is consumed alone. 




